---
title: Scalable Efficient Inference in Complex Surveys through Targeted Resampling of Weights
output: 
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Implementation of modeling framework proposed in the paper Das, S., Bandyopadhyay, D., and Pati, D. (2025+) "Scalable Efficient Inference in Complex Surveys through Targeted Resampling".

## Overview

Survey data often arises from complex sampling designs with unequal inclusion probabilities. When sampling is informative, traditional inference methods yield biased estimators and poor coverage. For efficient inference on survey data, we propose the *Survey-adjusted Weighted Likelihood Bootstrap* (S-WLB), which resamples weights from a from a carefully chosen distribution centered around the underlying sampling weights. S-WLB is computationally efficient, theoretically consistent, and delivers finite-sample uncertainty intervals which are proven to be asymptotically valid. 

## Functions

-  **normal_model.R** - Functions to get the estimators when data is generated from a normal model. Estimates are obtained from the following functions:

   - `SWLB.normal` - Our proposed Survey-adjusted Weighted Likelihood Bootstrap. 
   - `PMLE.normal` - Pseudo Maximum Likelihood Estimator. 
   - `WBB.normal` - Weighted Bayesian Bootstrap, Algorithm 2 of [Gunawan et al. (2020)](https://onlinelibrary.wiley.com/doi/abs/10.1111/anzs.12284). 
   - `BPPE.normal` - Bayesian Pseudo Posterior Estimator ([Savitsky \& Toth, 2016](https://projecteuclid.org/journals/electronic-journal-of-statistics/volume-10/issue-1/Bayesian-estimation-under-informative-sampling/10.1214/16-EJS1153.full)). 
   - `UBE.normal` - Unweighted Bayesian Estimator.
   

-  **probit_model.R** - Functions to get the estimators when data is generated from a probit regression model. Estimates are obtained from the following functions:

   - `SWLB.probit` - Our proposed Survey-adjusted Weighted Likelihood Bootstrap. 
   - `PMLE.probit` - Pseudo Maximum Likelihood Estimator. 
   - `WBB.probit` - Weighted Bayesian Bootstrap, Algorithm 3 of [Gunawan et al. (2020)](https://onlinelibrary.wiley.com/doi/abs/10.1111/anzs.12284). 
   - `BPPE.probit` - Bayesian Pseudo Posterior Estimator ([Savitsky \& Toth, 2016](https://projecteuclid.org/journals/electronic-journal-of-statistics/volume-10/issue-1/Bayesian-estimation-under-informative-sampling/10.1214/16-EJS1153.full)). 
   - `UBE.probit` - Unweighted Bayesian Estimator.
   
   
-  **data_generate.R** - Functions to generate data under a survey design with unequal probability of sampling. The functions `data_normal` and `data_probit` generate data from the normal model and the probit regression model, respectively. 
   

## Example

We create a population of size $N = 50,000$ where our random variable of interest $X$ and selection variable $Z$ jointly follow a bivariate normal distribution with means $\mu_x = 10$, $\mu_z = 0$, standard deviations $\sigma_x = 4$, $\sigma_z = 3$ and correlation $\rho = 0.5$.

For each population unit $l = 1, 2, \ldots, N$, the inclusion probabilities are generated as $\pi_l = \Phi(b_0 + b_1 Z_l)$, where $\Phi$ is the cumulative distribution function of a standard normal distribution, $b_0 = -1.8$ and $b_1 = 0.1$. 

Draw a sample of size $n = 500$ from the finite population with these inclusion probabilities. Let $\tilde{w}_{n,i} = n {\pi_i}^{-1} / \sum_{j=1}^n {\pi_j}^{-1}$ denote the scaled sampling weight of sampled unit $i$, where $\pi_i$ denotes its inclusion probability.

Our objective in this setting is to infer about the population mean, $\mu_x$.

**Data generation** :
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
source("data_generate.R")
source("normal_model.R")

set.seed(25)
# Specify the population and sample sizes
N = 50000; n = 500

# Specify the parameters to generate the data
mu_x = 10; mu_z = 0; s_x = 4; s_z = 3; rho = 0.5
b0 = -1.8; b1 =0.1;  M = 2000

# generate the data from the normal model
df = data_normal(mu_x = mu_x, mu_z = mu_z, s_x = s_x, s_z = s_z, rho = rho,
                 b0 = b0, b1 = b1, n = n, N = N)
```

**Get the point and interval estimates from the S-WLB method** :
```{r, message = FALSE, warning = FALSE}
est.SWLB = SWLB.normal(y = df$x, w = df$w, B = M)
est.SWLB
```

**Demonstrate the performance in point estimation and coverage probability of uncertainty intervals** :

We perform $100$ Monte Carlo simulations to compare the performance in point estimation and coverage probability of uncertainty intervals across the different methods.
```{r, message = FALSE, warning = FALSE}
# Specify the number of Monte Carlo simulations
nRep = 100
# Specify the methods for comparison
methods = c("PMLE", "SWLB", "BPPE", "UBE")

# Matrices to store the point estimates and coverage indicators
est = CovI = matrix(NA, nrow = nRep, ncol = length(methods))
colnames(est) = colnames(CovI) = methods

set.seed(25)
for(r in 1:nRep){
  
  # generate the data
  df = data_normal(mu_x = mu_x, mu_z = mu_z, s_x = s_x, s_z = s_z, rho = rho,
                   b0 = b0, b1 = b1, n = n, N = N)
  
  # PMLE
  est.PMLE = PMLE.normal(y = df$x, w = df$w)
  est[r, "PMLE"] = est.PMLE$mu.PMLE
  CovI[r, "PMLE"] = 1*(mu_x >= est.PMLE$CI.PMLE[1] && mu_x <= est.PMLE$CI.PMLE[2])
  
  # UBE
  est.UBE = UBE.normal(y = df$x)
  est[r, "UBE"] = est.UBE$mu.UBE
  CovI[r, "UBE"] = 1*(mu_x >= est.UBE$CI.UBE[1] && mu_x <= est.UBE$CI.UBE[2])
  
  # BPPE
  est.BPPE = BPPE.normal(y = df$x, w = df$w)
  est[r, "BPPE"] = est.BPPE$mu.BPPE
  CovI[r, "BPPE"] = 1*(mu_x >= est.BPPE$CI.BPPE[1] && mu_x <= est.BPPE$CI.BPPE[2])
  
  # SWLB
  est.SWLB = SWLB.normal(y = df$x, w = df$w, B = M)
  est[r, "SWLB"] = est.SWLB$mu.SWLB
  CovI[r, "SWLB"] = 1*(mu_x >= est.SWLB$CI.SWLB[1] && mu_x <= est.SWLB$CI.SWLB[2])
}
```

```{r, include=FALSE}
themegg = theme(
  # LABLES APPEARANCE
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA),
  plot.title = element_text(hjust = 0.5, size=14, face= "bold", colour= "black" ),
  axis.title.x = element_text(size=14, face="bold", colour = "black"),    
  axis.title.y = element_text(size=14, face="bold", colour = "black"),    
  axis.text.x = element_text(size=12, colour = "black"), 
  axis.text.y = element_text(size=12, colour = "black"),
  strip.text.x = element_text(size = 12, face="bold", colour = "black" ),
  strip.text.y = element_text(size = 12, face="bold", colour = "black"),
  strip.background =  element_rect(fill = "transparent",colour = NA),
  axis.line.x = element_line(color="black", linewidth = 0.2),
  axis.line.y = element_line(color="black", linewidth =  0.2),
  panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.2),
  legend.position = "none"
)
```

Plot showing the point estimates:

```{r, warning=FALSE, message=FALSE}
# all plots follow a pre-specified theme stored in "themegg", 
# which can be found in the README.Rmd file.

# data frame for plotting the point estimates
df.est = data.frame(est = c(est), method = rep(methods, each = nRep))
df.est$method = factor(df.est$method, levels = methods)

# Boxplots of the point estimates 
g.est = ggplot(df.est, aes(x = method, y = est, color = method)) +
  geom_boxplot(outlier.size = 0.2) + 
  geom_hline(yintercept = mu_x, linetype = "dotted", color = "darkblue") +
  scale_color_manual(values = c( "brown2", "forestgreen","orange", "grey30")) +
  xlab("Methods") + ylab("Estimate of the mean") + themegg
g.est
```

Plot showing the coverage probability of the $95$\% intervals:

```{r, warning=FALSE, message=FALSE}
# all plots follow a pre-specified theme stored in "themegg", 
# which can be found in the README.Rmd file.

# Calculate the coverage probability
CovP = colMeans(CovI)

# data frame for plotting the coverage probability
df.cov = data.frame(covP = CovP, method = methods)
df.cov$method = factor(df.cov$method, levels = methods)

# Barplots of the coverage probability
g.cov = ggplot(df.cov, aes(x = method, y = covP, color = method, fill = method)) +
  geom_bar(stat="identity", width = 0.5, alpha = 0.6)+
  geom_hline(yintercept = 0.95, linetype = "dotted", color = "darkblue") +
  scale_color_manual(values = c( "brown2", "forestgreen","orange", "grey30")) +
  scale_fill_manual(values = c( "brown2", "forestgreen","orange", "grey30")) +
  xlab("Methods") + ylab("Coverage probability") + themegg
g.cov
```